Demystifying the Computer Science \\Ph.D. Admission in US Universities\\{\large A guide for Vietnamese and International Students

### 2023-06-12 phd-cs-us.tex
Updates from Overleaf

\newcommand{\mycomment}[3][\color{black}]{{#1{{#2}: {#3}}}}
\newcommand{\tvn}[1]{\mycomment{TVN}{#1}}{}
\newcommand{\tl}[1]{\mycomment{ThanhLe}{#1}}{}
\newcommand{\red}[1]{{\color{red}{#1}}}
\renewcommand{\figurename}{Fig.}
\renewcommand{\tablename}{Tab.}
\title{Demystifying the Computer Science Ph.D. Admission in \\US Universities\\{\Large A guide for Vietnamese and International Students}}
\date{\today}
\author{\small \href{https://nguyenthanhvuh.github.io}{ThanhVu (Vu) Nguyen}, \href{https://cs.gmu.edu}{George Mason University}}
%\setcounter{secnumdepth}{3}
\begin{document}
\maketitle

@@

This document is on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{Github}. If you have questions or comments, feel free to create a \href{https://github.com/nguyenthanhvuh/phd-cs-us/issues}{GitHub issue} for discussion.
\end{abstract}

@@

\begin{wrapfigure}{r}{0.5\textwidth}
  \begin{center}
    %\includegraphics[width=0.48\textwidth]{birds}
  \end{center}
  \caption{Birds}
\end{wrapfigure}
While you might not have control over LORs \hyperref[sec:your-school]{where your go to school}, you do over your
statement! So write it well because we do take it seriously.
A well-written LOR also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \S\ref{sec:funding}).
%is important because if you need (GTA) funding, it will provide evidence
%that you can teach and communicate with students.

There are many guides on writing statement, e.g.,~\cite{blattman2022writing}, and\href{https://cs-sop.org/}{example statements},  so I will not talk too much about it. In short, discuss about your research vision and convince us that you can achieve it through your experience, e.g., published papers, or if you work on some projects by yourself, talk about it.

Finally, this is something easy to do, but is missed by many
applicants: \textbf{customize the statement} for the school you're applying to,
e.g., why do you apply here? talk about a couple of professors who you're interested in working with (in many cases your application will be forwarded to them for evaluation).
This shows that you're serious and have done homework on where you're applying to.
Be careful not to send wrong statement to wrong school or mixing
facts (e.g., talking about school X but mentioned about working with
profs. at school Y; and definitely do not talk about George Washington when applying to George Mason). I have seen such statements more time that I
should.

@@

\begin{itemize}[leftmargin=*]
\item International students not familiar with US universities often put them into \emph{two} bins:  (i) very top school that they dream about such as Stanford, MIT, Princeton, Harvard and (ii) everything else.  Sometimes they apply reputations of schools for non-CS programs, e.g., medical or physics, to rank CS programs. 
In some cases they rank universities based on popular states they know in the US, e.g., California, New York, and Texas.  Let's just say there are so many thing wrong with these methods.

@@

\item But more generally, rankings is superficial and you will want to do more research to be informed and make better decision. For example, if you get admissions to several places, you should contact profs. that you're interested in at those place and talk to them. They would be more willing to chat to you now that you have been admitted.  Ask them questions about their work, how they manage students, their expectations. You can even ask to contact their students.
\end{itemize}

    \item Can I apply to PhD to CS if my undergrad was not in CS or any related areas?
    \begin{itemize}[leftmargin=*]
        \item Yes, as long as you can demonstrate you are ready for CS PhD research through research experiences, LoRs, statements, etc as mentioned. You might be even able leverage this to make your profile stand out.
    \end{itemize}

>>>>

\newcommand{\mycomment}[3][\color{black}]{{#1{{#2}: {#3}}}}
\newcommand{\tvn}[1]{\mycomment{TVN}{#1}}{}
\newcommand{\tl}[1]{\mycomment{ThanhLe}{#1}}{}
\newcommand{\red}[1]{{\color{red}{#1}}}
\renewcommand{\figurename}{Fig.}
\renewcommand{\tablename}{Tab.}
\title{Demystifying the Computer Science \\Ph.D. Admission in US Universities\\{\large A guide for Vietnamese and International Students}}
\date{\today}
\author{\small \href{https://nguyenthanhvuh.github.io}{ThanhVu (Vu) Nguyen}, \href{https://cs.gmu.edu}{George Mason University}}
%\setcounter{secnumdepth}{3}
\begin{document}
\maketitle

@@

This document is available on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{Github}. If you have questions or comments, feel free to create a \href{https://github.com/nguyenthanhvuh/phd-cs-us/issues}{GitHub issue} for discussion.
\end{abstract}

@@

While you might not have control over LORs \hyperref[sec:your-school]{where your go to school}, you do over your
statement! So write it well because we do take it seriously.
A well-written LOR also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \S\ref{sec:funding}).
%is important because if you need (GTA) funding, it will provide evidence
%that you can teach and communicate with students.

There are various guides on writing statement, e.g.,~\cite{blattman2022writing}, and \href{https://cs-sop.org/}{many example statements},  so I will not talk too much about it. In short, discuss about your research vision and convince us that you can achieve it through your experience, e.g., published papers, or if you work on some projects by yourself, talk about it.

\begin{wrapfigure}{r}{0.47\textwidth}
\vspace{-0.4in}
  \begin{center}
    \includegraphics[width=0.47\textwidth]{c2.png}
  \end{center}
\vspace{-0.35in}  
  %\caption{Birds}
\end{wrapfigure}
Finally, this is something easy to do, but is missed by many
applicants: \textbf{customize the statement} for the school you're applying to,
e.g., why do you apply here? talk about a couple of professors who you're interested in working with (in many cases your application will be forwarded to them for evaluation).
This shows that you're serious and have done homework on places you're applying to.
Be careful not to send wrong statement to wrong school or mixing
facts (e.g., talking about school X but mentioned about working with
profs. at school Y; and definitely do not talk about George Washington when applying to George Mason). I have seen such statements more time that I
should.

@@

   \begin{center}
    \includegraphics[scale=0.5]{c1.png}
   \end{center}

\begin{itemize}[leftmargin=*]
\item  International students not familiar with US universities often put them into \emph{two} bins:  (i) very top school that they dream about such as Stanford, MIT, Princeton, Harvard and (ii) everything else.  Sometimes they rank CS programs using the reputations non-CS programs, e.g., medical or physics.
In some cases they rank universities based on popular states they know in the US, e.g., California, New York, and Texas.  Let's just say there are so many thing wrong with these methods.

@@

\item More generally, rankings can be superficial and you will want to do more research to be informed and make better decision. For example, if you get admissions to several places, you should contact profs. that you're interested in at those place and talk to them. They would be more willing to chat to you now that you have been admitted.  Ask them questions about their work, how they manage students, their expectations. You can even ask to contact their students.
\end{itemize}

    \item Can I apply to PhD in CS if my undergrad was not in CS or any related areas?
    \begin{itemize}[leftmargin=*]
        \item Yes, as long as you can demonstrate you are ready for CS PhD research through research experiences, LoRs, statements, etc as mentioned. You might be even able leverage this to make your profile stand out.
    \end{itemize}
